---
title: "Publishing Interim Research Products"
author: "Hao Ye"
institute: "Health Science Center Libraries, University of Florida"
date: "(updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: xaringan-themer.css
    lib_dir: libs
    seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      highlightSpan: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE)
```

```{r xaringan-themer, include=FALSE, warning=FALSE}
library(xaringanthemer)
style_mono_accent(
  base_color = "#09274B",
  base_font_size = "32px", 
  header_h1_font_size = "2.0rem",
  header_h2_font_size = "1.5rem",
  header_h3_font_size = "1.25rem", 
  header_background_content_padding_top = "2rem", 
  header_font_google = google_font("Lato"),
  text_font_google   = google_font("Montserrat", "300", "300i", "500"),
  code_font_google   = google_font("Source Code Pro"), 
  extra_css = list("p" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "ul" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "li" = list("margin-block-start" = "0.5em", 
                              "margin-block-end" = "0.5em"), 
                   "pre" = list("margin-top" = "0.5em", 
                              "margin-bottom" = "0.5em")), 
  link_color = "#339944"
)
```
class: inverse, center, middle

# `r rmarkdown::metadata$title`
### `r rmarkdown::metadata$author`
### `r rmarkdown::metadata$institute`
### (updated: `r Sys.Date()`)

<script async src="https://platform.twitter.com/widgets.js" charset="utf-8"></script>

---
# Motivations

Why should you care about this topic?

--

**Academic systems evaluate people**
* I dislike this, but that doesn't make it less true

--

Success requires self-advocacy

* NOT the same as boasting
* make it **easy for other people to find and learn about your work**

---
# Learning Outcomes

* understand the FAIR principles as they relate to scientific data management
* explain the benefits of ORCID and Google Scholar profiles
* define DOI, its attributes, and uses
* identify tools and platforms for publishing interim research products

---
class: inverse, center, middle

# FAIR

---
# FAIR

* Guidelines for *Scientific Data*
  (but which apply more generally)
  - Findable
  - Accessible
  - Interoperable
  - Reusable

---
# Findable

* Can humans and machines find the data?
  - use metadata that accurately describe the content
  - use metadata that helps other people find the content
  
---
# Metadata

data that describes or gives information about other data

--
<blockquote class="twitter-tweet"><p lang="en" dir="ltr">The ISO country code for Namibia is NA, which has been a hilariously frustrating discovery I made while coding today <a href="https://twitter.com/hashtag/rstats?src=hash&amp;ref_src=twsrc%5Etfw">#rstats</a></p>&mdash; Sarah Bowden (@mamabphd) <a href="https://twitter.com/mamabphd/status/1273352107240968193?ref_src=twsrc%5Etfw">June 17, 2020</a></blockquote> 

* hashtag identifies content as related to R
---
# Accessible

* there should be minimal barriers to access the data
  - e.g. no proprietary interfaces or software required
* use established methods to authenticate access to protected or private data

---
# Interoperable

* use a common, shared format for storage
* use a metadata standard to document attributes
  - [FAIR Data Point](https://github.com/FAIRDataTeam/FAIRDataPoint/wiki/FAIR-Data-Point-Specification) is one
  - some are domain-specific, e.g. [Ecological Metadata Language](https://eml.ecoinformatics.org/)

---
# Reusable

* describe the data and metadata well for researchers to identify whether it is suitable for their usage
* use a legal license that enables re-use
  - e.g. CC-BY or CC-0
  - **don't use a code license**, e.g. MIT
* include attribution and a description of how to cite the data

---
# Summary

* FAIR principles are intended to make data more useful
  - the same principles apply for academic work!
  
* for more on data, see the [Research Data Management Libguide](https://guides.uflib.ufl.edu/datamanagement) 

---
class: inverse, center, middle

# Personal Profiles

---
# Your Academic Identity

<blockquote class="twitter-tweet"><p lang="en" dir="ltr">Never ceases to amaze me how many postdoc scientists &amp; other early career researchers have basically zero internet presence. Many have no website (beyond an impoverished uni site &amp; even that&#39;s often missing), no readily available contact details, no pub list.<br><br>Don&#39;t be invisible.</p>&mdash; Chris Chambers (@chrisdc77) <a href="https://twitter.com/chrisdc77/status/1090198315617132545?ref_src=twsrc%5Etfw">January 29, 2019</a></blockquote>

---
# is it FAIR?

* Findable
* Accessible
* Interoperable
* Reusable

---
# ORCID

* https://orcid.org/
* Open Researcher and Contributor IDentifier
* free to use
* unique ID that stays with you, regardless of changes in name, email, institution
* common way to login to journals and identify yourself on publications

---
# Google Scholar

* https://scholar.google.com/
* free to use
* track your works, correct issues in them
* citation statistics (imperfect, but easy for people to check)

---
# Other

* LinkedIn - if you want to make industry connections
* whatever other sites are commonly used in your field
* a personal website:
  - lots of options here
  - let me know if a workshop on making websites would be useful!
  
---

class: inverse, center, middle

# DOI

---
# DOI = Digital Object Identifier

* format
  - `10.NNNN/{suffix}`
  - `NNNN` identifies the registrant
  - `{suffix}` identifies the object
  - e.g. `10.5281/zenodo.3892184`
---
# DOI Properties

* DOIs are permanent
  - the object and metadata record are archived
* unambiguous
  - each DOI links to a unique object
  
---
# Why is this Important?

A given DOI always point to the same object.
- (some exceptions for updating the metadata or deleting the object)

The scholarly record needs permanence.
- someone else needs to be able to look up your sources
- DOIs fit this role

Most journals issue a DOI for each article.

---
# Get DOIs for Your Work

* If you have interim products, e.g.
  + code
  + datasets
  + protocols
  + images/video
--

* **A DOI makes that work citable!**
  + (for fellowship/grant/job applications)

---
class: inverse, center, middle

# Publishing Platforms

---
# Places that issue DOIs

* preprint servers
* data repositories
* general-purpose repositories
* tooling-specific repositories

---
# Preprints

* places to post papers *pre*-publication
* very common in math & physics, catching on elsewhere

--

why?

* share what your work more quickly
  - not delayed until peer review
* (sometimes) get feedback

---
# Preprint servers

* arXiv -- primary one for math and physics
  - no DOI
* bioRxiv -- primary one for biology
* medRxiv -- primary one for health sciences
* many others
* most are free

---
# Data Repositories

* Zenodo
  - https://zenodo.org/
  - approved by UF IT for data storage (workshops forthcoming)
  - free to use, 50 GB capacity, supported by CERN
* Dryad - https://datadryad.org
* figshare - https://figshare.com

---
# Data Repositories

Additional Notes
* Use whatever is common for your field / journal
* Be aware of sharing restrictions for protected/private data
* [Research Data Management Libguide](https://guides.uflib.ufl.edu/datamanagement) 

---
# General Purpose Repositories

* Open Science Framework
  - https://osf.io/
  - files can be uploaded or linked through github, dropbox, google drive, etc.
  - can embed projects within other projects
  - can issue DOIs for public projects

---
# Tooling Specific Repositories

For specific (common) applications:
* laboratory protocols and handy app -- https://www.protocols.io/
* computer code -- https://github.com
  - does not issue DOI, but deposits to Zenodo
* pre-register study design and analysis -- https://clinicaltrials.gov/, or https://osf.io/
* [list of journals with registered reports](https://www.cos.io/our-services/registered-reports)

---

# Thanks

* Let me know what content you'd like to see
* Contact me for additional questions or consultation requests!
* Check back in on the libguide for more modules and contact info:
  - https://guides.uflib.ufl.edu/reproducibility

